Forecasting the 2024 U.S. Presidential Election

### 2024-11-03 paper/paper.qmd
Revise references, llm usages

This paper focuses on forecasting the electoral outcomes between Kamala Harris and Donald Trump in the 2024 cycle. Using the multiple linear regression model build on the polling data from @fivethirty, we construct an analysis by incorporating factors such as poll quality, transparency, sample size, state preference, and poll timing. By integrating these variables, the model produces a detailed forecast that accounts for regional differences and methodological variations, offering a thorough view of projected support for each candidate. The result indicates a competitive election landscape, with Harris showing strong prospects in key swing states, while Trump retains support in traditional Republican states. These findings demonstrate how the persistent influence of geographic voting patterns and the polling characteristics would influence the electoral votes.

This paper offers valuable analysis for political analysts, campaign strategists, and the public, enhancing the understanding of how polling data can inform electoral outcome predictions. However, recent election cycles have also revealed limitations in polling accuracy due to biases, non-response errors, and variations in poll quality. This study proposes a feasible model that addresses these issues to improve the reliability of election forecasts, and the uncertainty beyond the data has also been acknowledged in this paper.
The paper is organized as follows: @sec-data describes the dataset and the processing steps for model preparation. @sec-modelsetup explains the modeling approach, including predictor selection and the regression model structure. @sec-results presents the primary findings, detailing state-by-state predictions and overall electoral projections. @sec-discussion discusses the key findings, regional voting patterns, polling uncertainty and weakness. The @sec-Appendix1 provides the evaluation of the survey USA and @sec-Appendix2 discusses about the idealized methodology on survey. Besides, model diagnostics and summary are also attached in @sec-model.

The estimand in this paper represents the actual influence of certain polling factors on the forecasted support for Kamala Harris and Donald Trump in the 2024 election. Specifically, it examines how factors like poll quality, transparency, sample size, and timing impact voting percentages for each candidate across different states. Defining this estimand here provides a consistent foundation for the analysis, ensuring that the model accurately reflects the relationship between these polling characteristics and candidate support while minimizing potential distortions from methodological differences.

@@

The outcome variable of interest for this research is the percentage representing the level of public support for Donald Trump. The distribution is shown in @fig-pct, where support percentage for Donald Trump ranges from approximately 20% to 70%. The red dashed line at the 50% mark acts as a benchmark, indicating the threshold needed for majority support.

@@

The numeric grade reflects the quality of the pollster, with FiveThirtyEight defining a scale from 0 to 3. A grade of 0 indicates a low-quality poll, while a grade of 3 signifies a high-quality pollster. After filtering for pollsters with a numeric grade higher than 2.5, we identified a total of 30 distinct pollsters. The distribution shows that grades cluster around 2.6, 2.8, and 2.9, with fewer polls receiving grades of 2.7 and 3.0, as shown in @fig-numeric-grade .

@@

The Transparency Score measures how transparent a pollster is, calculated based on the amount of information disclosed about its polls, weighted by recency. The highest possible score is 10, while the lowest is 0. The distribution of Transparency Scores for the filtered pollsters shows a peak around 9 as presented in @fig-transparency-score, indicating that nine the most common transparency score. @fig-transparency-score suggests that among the selected pollsters, there is a predominance of high transparency scores.

@@

```{r}
#| label: fig-end-date
#| fig-cap: Trend of Support Percentage for Donald Trump Over Time Leading Up to the 2024 Election Day. The red dashed line represents the smoothed trend of average support for Donald Trump at each point in time, showing fluctuations and a gradual increase in support as the election day approaches.Renew the data to the October 24.2024
#| echo: false
#| warning: false
#| message: false
ggplot(analysis_data, aes(x = days_until_election, y = percent)) + 
    geom_point(color = "skyblue", size = 1) +  # Smaller dots
    geom_smooth(method = "loess", color = "red", linetype = "dashed", size = 1) + 
    labs(
        x = "Days Remaining Until Election Day", 
        y = "Support Percentage for Donald Trump (%)"
    ) + 
    theme_minimal(base_size = 14) + 
    scale_x_reverse() + 
    theme(
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        axis.text = element_text(color = "#555555"),
        panel.grid.minor = element_blank()
    )

@@

Additional details and model diagnostics are provided in [Appendix -@sec-model-details].

@@

The rationale for applying multiple regression in this context is to control for various influential factors simultaneously—such as time trends (`days_until_election`), pollster quality (`transparency_score`), and regional differences (`state-level effects`). This approach allows us to isolate the individual impact of each variable, providing a more detailed understanding of how each factor contributes to Trump’s polling outcomes. The full coefficient output is shown in the [Appendix -@sec-model-details]

@@

The table below (@tbl-prediction) summarizes the predicted results, showing Trump’s predicted percentage, the number of electoral votes in each state, and the predicted winner (either Trump or Harris). For instance:
- Alabama: Trump is predicted to win 53% of the vote, securing all 9 electoral votes.
- California: Trump is predicted to receive 34.39% of the vote, resulting in a victory for Harris, who takes California's 55 electoral votes.
- Florida: The model predicts a close race, with Trump at 48.89% of the vote, resulting in a Harris win in this battleground state.

@@

## Predicted Electoral Outcomes by State
The table in the [Appendix -@tbl-missing-states] supplements the predicted electoral outcomes by providing likely results for states where the model could not generate predictions. Using historical data and winner information to estimate outcomes for these missing states ensures a complete view of electoral projections and accounts for every state in the final electoral vote totals.

@@

@tbl-numstate shows the numerical predicted winner of the 2024 US Presidential Election by state. Harris is expected to win the 2024 U.S. Presidential Election, securing 387 electoral votes, compared to Trump’s 151 electoral votes. The election outcome hinges on several battleground states, where the vote margins are predicted to be narrow.
```{r}
#| echo: false
#| label: tbl-numstate
#| tbl-cap: Numerical Projected Electoral Vote Count for the 2024 U.S. Presidential Election (Based on Polling Data and Historical Trends). 
#| warning: false
#| message: false

@@

The analysis shows that state-specific factors, such as historical voting patterns and days until election day, significantly influence each candidate’s polling outcomes. Traditional Democratic strongholds like California and New York favor Harris, while Republican-leaning states such as Alabama and Wyoming align with Trump. This distribution reflects persistent geographic voting patterns and highlights the impact of regional party loyalty.

@@

## Regional Voting Patterns and Their Impact on Electoral Outcomes
Our analysis demonstrates the persistence of geographic voting patterns in U.S. elections. As expected, traditional Democratic strongholds such as California, New York, and Illinois favor Harris, while Republican-dominated states like Alabama and Wyoming show support for Trump. These regional preferences align with historical voting behaviors, suggesting that party loyalty and demographic factors continue to shape election results.

The model also highlights the significant role of battleground states, with states such as Florida and Arizona predicted to lean toward Harris. The influence of these states, where margins are often narrow, emphasizes the importance of campaign efforts and voter mobilization in these regions. The projections illustrate that minor shifts in voter sentiment within these states could decisively impact the overall election outcome because of the high number of electoral votes.

@@

# Pollster Methodology Overview and Evaluation {#sec-Appendix1}

@@

* **Advantages:** The diverse sampling approach not only ensures a broad range of opinions is captured but also complements probability-based sampling, which accurately reflects the overall population. Furthermore, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Additionally, rotating questions and answer choices helps mitigate bias, further improving the reliability of the data. Finally, the use of online surveys offers a cost-effective solution for efficient data collection.\

* **Disadvantages:** Phone-based data collection tends to be time-consuming and can be affected by interviewer effects during telephone interviews. Additionally, challenges like non-response issues, such as busy signals or refusals to participate, can hinder the effectiveness of the data collection process.\

@@

Non-response in survey research is important to address because it introduces potential bias, impacting the accuracy and representativeness of the results.As mentioned by @groves2012nonresponse, when certain groups are underrepresented due to non-response, it can skew findings, making them less reflective of the actual population. For example, individuals who with demanding schedules, lower income or limited access to technology might be less likely to respond to a poll. Meanwhile, people with strong opinions or from specific demographic groups could be overrepresented in the results. As a result of non-response, the findings may misinform decision-makers, leading to conclusions that don’t truly represent the diverse views of the entire population. Addressing non-response is thus essential for producing balanced, reliable survey results that can genuinely inform policies, programs, and public opinion.\

@@

* **Positive Aspects:**
A logical flow between questions facilitates easy navigation for respondents throughout the survey, while simple wording promotes inclusivity by enabling individuals from diverse backgrounds to comprehend the questions. Furthermore, all questions are directly relevant to analyzing the 2024 U.S. presidential election, and providing predefined response options simplifies the choices for participants.\
* **Negative Aspects:**
Static options for party affiliation and ideology may overlook the subtleties of respondents' political beliefs. These fixed categories risk oversimplifying complex political identities.\

@@

# Idealized Methodology and Survey {#sec-Appendix2}

@@

### Sample Size
A total of **10,000 respondents** will be surveyed, providing a margin of error of approximately ±1\% at a 95\% confidence level. This sample size allows for detailed subgroup analysis (e.g., by state or demographic group), yielding statistically robust predictions that represent the target population.

### Weighting
Post-stratification weights will be applied to adjust for any oversampling or undersampling of specific demographic groups. For instance, younger voters or underrepresented minorities will be weighted to reflect their true proportions in the voting population. This weighting ensures that the final sample aligns closely with the demographic makeup of the target population.

## Recruitment Strategy
Recruitment is executed through **multi-channel outreach**, combining digital advertisements, email outreach, and partnerships with civic organizations. This strategy is designed to capture a probability sample and follows best practices in survey methodologies to minimize non-response bias—errors that occur when certain groups in the target population are less likely to respond to the survey. Non-response matters because it can distort the results if certain demographics are underrepresented in the final data [@pewresearch].

@@

# Model details {#sec-model}

>>>>

This paper focuses on forecasting the electoral outcomes between Kamala Harris and Donald Trump in the 2024 cycle. Using the multiple linear regression model build on the polling data from @fivethirty, we construct an analysis by incorporating factors such as poll quality, transparency, sample size, state preference, and poll timing. By integrating these variables, the model produces a detailed forecast that accounts for regional differences and methodological variations, offering a thorough view of projected support for each candidate. The results indicate a decisive victory for Harris, who is projected to secure approximately 387 electoral votes, showcasing strong performance in key swing states and surpassing the required 270 votes by a wide margin. In contrast, Trump is predicted to receive 151 electoral votes, retaining support primarily within traditional Republican strongholds. These findings underscore Harris's substantial lead and emphasize the ongoing impact of geographic voting patterns and polling characteristics on electoral outcomes.

This paper offers valuable analysis for political analysts, campaign strategists, and the public, enhancing the understanding of how polling data can inform electoral outcome predictions. However, recent election cycles have also revealed limitations in polling accuracy due to biases, non-response errors, and variations in poll quality. This study proposes a feasible model that addresses these issues to improve the reliability of election forecasts, and the uncertainty beyond the data has also been acknowledged in this paper.
The paper is organized as follows: @sec-data describes the dataset and the processing steps for model preparation. @sec-modelsetup explains the modeling approach, including predictor selection and the regression model structure. @sec-results presents the primary findings, detailing state-by-state predictions and overall electoral projections. @sec-discussion discusses the key findings, regional voting patterns, polling uncertainty and weakness. The @sec-Appendix1 provides n evaluation of the SurveyUSA polls and @sec-Appendix2 discusses about the idealized methodology on survey. Besides, model diagnostics and summary are also attached in @sec-model.

The estimand in this paper represents the actual influence of certain polling factors on the forecasted support for Kamala Harris and Donald Trump in the 2024 election. Specifically, it examines how factors like poll quality, transparency, sample size, and timing impact voting percentages for each candidate across different states. However, the true estimand remains unknown, as we can only approximate it using available polling data and model assumptions. Defining this estimand here provides a consistent foundation for the analysis, ensuring that the model accurately reflects the relationship between these polling characteristics and candidate support while minimizing potential distortions from methodological differences.

@@

The outcome variable of interest for this research is the percentage representing the level of public support for Donald Trump. The distribution is shown in @fig-pct, where support percentage for Donald Trump ranges from approximately 20% to 70%. The red dashed line at the 50% mark acts as a benchmark, indicating the threshold needed for majority support for winning.

@@

The numeric grade reflects the quality of the pollster, with @fivethirty defining a scale from 0 to 3. A grade of 0 indicates a low-quality poll, while a grade of 3 signifies a high-quality pollster. After filtering for pollsters with a numeric grade higher than 2.5, we identified a total of 30 distinct pollsters. The distribution shows that grades cluster around 2.6, 2.8, and 2.9, with fewer polls receiving grades of 2.7 and 3.0, as shown in @fig-numeric-grade .

@@

The Transparency Score measures how transparent a pollster is, calculated based on the amount of information disclosed about its polls, weighted by recency. The highest possible score is 10, while the lowest is 0. The distribution of Transparency Scores for the filtered pollsters shows a peak around 9 as presented in @fig-transparency-score, indicating that nine is the most common transparency score. @fig-transparency-score suggests that among the selected pollsters, there is a predominance of high transparency scores.

@@

```{r}
#| label: fig-end-date
#| fig-cap: Trend of Support Percentage for Donald Trump Over Time Leading Up to the 2024 Election Day. The red dashed line represents the smoothed trend of average support for Donald Trump at each point in time, showing fluctuations and a gradual increase in support as the election day approaches. Renew the data to the October 24, 2024. 
#| echo: false
#| warning: false
#| message: false
ggplot(analysis_data, aes(x = days_until_election, y = percent)) + 
    geom_point(color = "skyblue", size = 1) +  # Smaller dots
    geom_smooth(method = "loess", color = "red", linetype = "dashed", size = 1) + 
    labs(
        x = "Days Remaining Until Election Day", 
        y = "Support Percentage for Donald Trump (%)"
    ) + 
    theme_minimal(base_size = 14) + 
    scale_x_reverse() + 
    theme(
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        axis.text = element_text(color = "#555555"),
        panel.grid.minor = element_blank()
    )

@@

Additional details and model diagnostics are provided in @sec-model.

@@

The rationale for applying multiple regression in this context is to control for various influential factors simultaneously—such as time trends (`days_until_election`), pollster quality (`transparency_score`), and regional differences (`state-level effects`). This approach allows us to isolate the individual impact of each variable, providing a more detailed understanding of how each factor contributes to Trump’s polling outcomes. The full coefficient output is shown in the @sec-model

@@

The table below (@tbl-prediction) summarizes the predicted results, showing Trump’s predicted percentage, the number of electoral votes in each state, and the predicted winner (either Trump or Harris). For instance:

- South Dakota: Trump is predicted to win 57.17% of the vote, securing all 3 electoral votes for Trump.
- California: Trump is predicted to receive 34.39% of the vote, resulting in a victory for Harris, who takes California's 55 electoral votes.
- Florida: The model predicts a close race, with Trump at 50.04% of the vote, resulting in a Trump win in this battleground state.

@@

## Predicted Electoral Outcomes by State
The table in the @tbl-missing-states supplements the predicted electoral outcomes by providing likely results for states where the model could not generate predictions. Using historical data and winner information to estimate outcomes for these missing states ensures a complete view of electoral projections and accounts for every state in the final electoral vote totals.

@@

@tbl-numstate shows the numerical predicted winner of the 2024 US Presidential Election by state. Harris is expected to win the 2024 U.S. Presidential Election, securing 387 electoral votes in total, compared to Trump’s 151 electoral votes. The election outcome hinges on several battleground states, where the vote margins are predicted to be narrow.
```{r}
#| echo: false
#| label: tbl-numstate
#| tbl-cap: Numerical Projected Electoral Vote Count for the 2024 U.S. Presidential Election (Based on Polling Data and Historical Trends). 
#| warning: false
#| message: false

@@

The analysis shows that state-specific factors, such as historical voting patterns and days until election day, significantly influence each candidate’s polling outcomes. Traditional Democratic strongholds like California and New York favor Harris, while Republican-leaning states such as South Dakota and Missouri align with Trump. This distribution reflects persistent geographic voting patterns and highlights the impact of regional party loyalty.

@@

## Regional Voting Patterns and Their Impact on Electoral Outcomes
Our analysis demonstrates the persistence of geographic voting patterns in U.S. elections. As expected, traditional Democratic strongholds such as California, New York, and Massachusetts favor Harris, while Republican-dominated states like South Dakota and Missouri show support for Trump. These regional preferences align with historical voting behaviors, suggesting that party loyalty and demographic factors continue to shape election results.

For instance, Florida, with its 30 electoral votes, is projected to lean toward Trump, with 50.04% of voters supporting him. In contrast, Arizona, with its 11 electoral votes, is expected to lean toward Harris, as 47.64% of voters are predicted to support Trump. The influence of these states, where margins are often narrow, emphasizes the importance of campaign efforts and voter mobilization in these regions. The projections illustrate that minor shifts in voter sentiment within these states could decisively impact the overall election outcome because of the high number of electoral votes.

@@

# Appendix: Pollster Methodology Overview and Evaluation {#sec-Appendix1}

@@

* **Advantages:** The diverse sampling approach not only ensures a broad range of opinions is captured but also complements probability-based sampling, which accurately reflects the overall population. Furthermore, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Additionally, rotating questions and answer choices helps mitigate bias, further improving the reliability of the data. Finally, the use of online surveys offers a cost-effective solution for efficient data collection.\
* **Disadvantages:** Phone-based data collection tends to be time-consuming and can be affected by interviewer effects during telephone interviews. Additionally, challenges like non-response issues, such as busy signals or refusals to participate, can hinder the effectiveness of the data collection process.

@@

Addressing non-response in survey research is essential because it introduces potential bias, particularly non-response bias, which can impact the accuracy and representativeness of the results. As mentioned by @groves2012nonresponse, when certain groups are underrepresented due to non-response, it can skew findings, making them less reflective of the actual population. For example, individuals who with demanding schedules, lower income or limited access to technology might be less likely to respond to a poll. Meanwhile, people with strong opinions or from specific demographic groups could be overrepresented in the results. As a result of non-response, the findings may misinform decision-makers, leading to conclusions that don’t truly represent the diverse views of the entire population. Addressing non-response is thus essential for producing balanced, reliable survey results that can genuinely inform policies, programs, and public opinion.

@@

* **Positive Aspects:**
A logical flow between questions facilitates easy navigation for respondents throughout the survey, while simple wording promotes inclusivity by enabling individuals from diverse backgrounds to comprehend the questions. Furthermore, all questions are directly relevant to analyzing the 2024 U.S. presidential election, and providing predefined response options simplifies the choices for participants.\
* **Negative Aspects:**
Static options for party affiliation and ideology may overlook the subtleties of respondents' political beliefs. These fixed categories risk oversimplifying complex political identities.

@@

# Appendix: Idealized Methodology and Survey {#sec-Appendix2}

@@

### Sample Size
Sample size is a critical factor in research design, as it influences the accuracy, reliability, and generalizability of study results. As mentioned by @cohen2013statistical, in statistical analysis, a larger sample size generally provides a more accurate reflection of the population characteristics, reducing the potential for sampling error.

Thus, for the idealized methodology, a total of **10,000 respondents** will be surveyed, providing a margin of error of approximately ±1\% at a 95\% confidence level. This sample size allows for detailed subgroup analysis (e.g., by state or demographic group), yielding statistically robust predictions that represent the target population.

### Weighting
As discussed by @kalton2003weighting, weighting adjustments are frequently applied in surveys to address non- response and non-coverage, helping weighted sample estimates align with known population values. Thus, in this idealized methodology, post-stratification weights will be used to adjust for any oversampling or undersampling of specific demographic groups. For example, younger voters or underrepresented minorities will be weighted to reflect their actual proportions in the voting population. This adjustment ensures that the final sample closely mirrors the demographic composition of the target population, enhancing the accuracy and representativeness of the results.

## Recruitment Strategy
Recruitment is executed through **multi-channel outreach**, combining digital advertisements, email outreach, and partnerships with civic organizations. This strategy is designed to capture a probability sample and follows best practices in survey methodologies to minimize non-response bias—errors that occur when certain groups in the target population are less likely to respond to the survey. Non-response matters because it can distort the results if certain demographics are underrepresented in the final data [@pewresearch]. Moreover, utilizing multiple methods can reduce costs, speed up response collection, and improve response rates [@dillman2014internet].

@@

# Appendix: Model details {#sec-model}
